---
title: "Data Exploration"
author: "François Delavy"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(rio)
library(glue)
library(lubridate)
library(stringr)

subm <- rio::import("vfsg_submissions.csv") %>% 
  mutate(ProjectDate = format(lubridate::parse_date_time(`Date of project`, orders = c("my")), "%Y-%m"),
         Volunteer = as.factor(`Name of volunteer`),
         URL = ifelse(stringr::str_detect(string = URL, pattern = "http"), URL, NA),
         `Volunteer City` = ifelse(stringr::str_detect(string = `Volunteer City`, pattern = "nknown"), NA, `Volunteer City`)
         )

print("TO DO: consolidate city names. E.g. Washington D.C., London, , Kawasaki-shi")


```

## Data

```{r}
glimpse(subm)
```

## Basic Stats

```{r}

glue("Number of observations: {dim(subm)[1]}")
glue("Number of projects: {subm$`Name of charity/Project` %>% unique() %>% length()}")
glue("Number of charity countries: {subm$`Charity Country` %>% unique() %>% length()}")
glue("Number of charity cities: {subm$`Charity City` %>% unique() %>% length()}")
glue("Number of topics: {subm$Topic %>% unique() %>% length()}")
glue("Projects range from {min(subm$ProjectDate)} to {max(subm$ProjectDate)}")
glue("Number of volunteers who participated {subm$`Name of volunteer` %>% unique() %>% length()}")
glue("Number of volunteer countries: {subm$`Volunteer Country`  %>% unique() %>% length()}")
glue("Number of volunteer cities: {subm$`Volunteer City`  %>% unique() %>% length()}")
glue("Number of unkown URLs: {subm$URL %>% is.na() %>% sum()}, that is {round(subm$URL %>% is.na() %>% sum() / subm$URL %>% is.na() %>% length() * 100, 1)}%")

```

## Basic plots

```{r}
toplot <- dplyr::count(subm, Volunteer, sort=TRUE) %>% head(20) %>% mutate(Volunteer = fct_reorder(Volunteer, n))
ggplot(toplot, aes(x = Volunteer, y = n)) +
  geom_col() + coord_flip() + theme_minimal() +
  ggtitle("Top 20 volunteers:")
```

```{r}
toplot <- subm %>% count(`Volunteer Country`, sort=TRUE)  %>% mutate(Volcount  = fct_reorder(`Volunteer Country`, n))
ggplot(toplot, aes(x = Volcount, y = n)) +
  geom_col() + coord_flip() + theme_minimal() +
  ggtitle("Volunteer's countries (not done by distinct volunteer)")
```

```{r, fig.height=18}
toplot <- subm %>% count(`Volunteer City`, sort = TRUE)  %>% mutate(Volcity  = fct_reorder(`Volunteer City`, n))
ggplot(toplot, aes(x = Volcity, y = n)) +
  geom_col() + coord_flip() + theme_minimal() +
  ggtitle("Volunteer's cities (not done by distinct volunteer)")
```


```{r}
toplot <- subm %>% count(`Charity Country`, sort=TRUE)  %>% mutate(Charcount  = fct_reorder(`Charity Country`, n))
ggplot(toplot, aes(x = Charcount, y = n)) +
  geom_col() + coord_flip() + theme_minimal() +
  ggtitle("Charity's countries (not done by distinct charity)")
```




```{r}
toplot <- subm %>% count(`Charity City`, sort=TRUE)  %>% mutate(Charcity  = fct_reorder(`Charity City`, n))
ggplot(toplot, aes(x = Charcity, y = n)) +
  geom_col() + coord_flip() + theme_minimal() +
  ggtitle("Charity's cities (not done by distinct charity)")
```


## Brainstorming

### Network
- on a map?
- very light map or no map plotted but you recognize the shape?
- node = participants_city (or participant?), link = project, size = n_participants from city (or n_project?)
- add the city of the charity on another city on the map
- facet by projects (the network of each project), or all in one? 

__Examples__  

- https://stackoverflow.com/questions/33122456/r-creating-a-world-network-map
- https://kateto.net/sunbelt2019 
- https://www.r-bloggers.com/2018/05/three-ways-of-visualizing-a-graph-on-a-map/
- https://mran.revolutionanalytics.com/snapshot/2016-03-14/web/packages/GGally/vignettes/ggnetworkmap.html (old?)
- https://rpubs.com/BrendanKnapp/GeospatialNetworkPlotting
- 

### Other
- time serie of number of participants
- 

## To Keep in Mind
- the goals (incite new volunteers or encourage charities to viz data)
- show impact with a few citation?
- have one clear message for the viz
- 



## Let's try to build a network for one project


```{r}

awb <- subm %>% 
  filter(`Name of charity/Project` == 'Academics Without Borders') %>%
  select(City = `Volunteer City`) %>%
  drop_na(City)

nodes <- awb$City %>% unique() %>% as_factor()

city_count <- awb %>% group_by(City) %>% count() # unique()

edges <- awb %>%
  mutate(Destination = paste(nodes, collapse=',')) %>%
  separate_rows(Destination) %>%
  unique() %>%
  left_join(city_count, by = 'City') %>%
  # mutate(from = as_factor(City), to = as_factor(Destination)) %>%
  select(from = City, to = Destination, weights = n)

# basic network:
library(tidygraph)
library(ggraph)

tibble()
graph <- as_tbl_graph(edges)

# test <- make_graph(edges$from) # awb$City)
df_test <- tibble()
test <- graph_from_data_frame(edges)

ggraph(graph, layout = 'graphopt', ) + 
  geom_edge_link() + 
  geom_node_point(size = 4) + 
  theme_graph()

```


```{r}
# geolocalize the cities:
library(maps)

world_cities <- maps::world.cities %>% rename(City = name)

vol_cities <- subm %>% 
  filter(`Name of charity/Project` == 'Academics Without Borders') %>%
  select(Country = `Volunteer Country`, City = `Volunteer City`) %>%
  drop_na(City) %>%
  left_join(world_cities, by = 'City')


world_map <- map_data("world")




ggplot(world_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") + 
  theme_bw()

```


